---
title: "Bearings Target Tracking"
author: "Bassel MASRI - Guillaume FRANCHI"
date: "11/17/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# INTRODUCTION

The aim of this project is to implement an approach to recover an estimation of the target's trajectory $x_{1:n}$ from the available corrupted observations $\theta_{1:n}^i$, $i=\{1,2\}$.

# SAMPLING DATA IN THE CASE OF ONE SENSOR

For this part, we consider that only one sensor works and its coordinates are $(s_x^1,s_y^1) = (-1.5,0.5)$ 

We will sample data $\theta_{1:n}^1$ according to the formula given below: 

$$
\theta_k^i = tan^-1(\frac{y_k-s_y^i}{x_k-s_x^i})+r_k \text{ with } r_k \sim N(0,\sigma^2)
$$

In order to do so, we first define the state space equations for the position and the velocity from the discrete Wiener velocity model (Nearly Constant Velocity model). The noise $q_k \sim N(0,Q)$ with $Q$ being the variance matrix defined.

$$
\begin{cases}
x_k = x_{k-1} + \Delta \dot{x}_{k-1} \\
y_k = y_{k-1} + \Delta \dot{y}_{k-1} \\
\dot{x}_k = \dot{x}_{k-1} \\
\dot{y}_k = \dot{y}_{k-1}
\end{cases}
+ q_k \sim N\left[\left(\begin{array}{c}
0\\
0\\
0\\
0
\end{array}\right),\left(\begin{array}{cccc}
q_1^c \frac{\Delta^3}{3} & 0 & q_1^c \frac{\Delta^2}{2} & 0\\
0 & q_2^c \frac{\Delta^3}{3} & 0 & q_2^c \frac{\Delta^2}{2}\\
q_1^c \frac{\Delta^2}{2} & 0 & q_1^c \Delta & 0 \\
0 & q_2^c \frac{\Delta^2}{2} & 0 & q_2^c \Delta
\end{array}\right)\right]
$$

The initial state to be considered is x$_1 = (0,0,1,0)^T$ which means that $x_0 = y_0 = \dot{y}_0 = 0$ and $\dot{x}_0 = 1$

We can proceed to implementing the state space model using the parameters described : 

```{r, warning=FALSE}
rm(list=ls())
library(MASS)

#-- Setting parameters
delta <- 0.01
q_1 <- q_2 <- 0.1
Q <- matrix(c((q_1*delta**3/3),0,(q_1*delta**2/2),0,
              0,(q_2*delta**3/3),0,(q_2*delta**2/2), 
              (q_1*delta**2/2),0,q_1*delta,0,
              0,(q_2*delta**2/2),0, q_2*delta), nrow = 4, byrow = TRUE)

#-- Define multivariate Gaussian noise of the state space model
k <- 500
q_k <- t(mvrnorm(k, rep(0,4), sqrt(Q)))
x <- x_dot <- y <- y_dot <- rep(0,k)

for (i in 1:k){
  
  #-- Initial values
  if (i==1){
    x[i] <- 0
    y[i] <- 0
    y_dot[i] <- 0
    x_dot[i] <- 1
  }
  
  #-- Implementing state space model and adding noise
  #-- note we add each element of the noise matrix to
  #-- its corresponding value.
  else {
    x[i] <- x[i-1] + delta*x_dot[i-1] + q_k[1,i]
    y[i] <- y[i-1] + delta*y_dot[i-1] + q_k[2,i]
    x_dot[i] <- x_dot[i-1] + q_k[3,i]
    y_dot[i] <- y_dot[i-1] + q_k[4,i]
  }
}
```


```{r, fig.align = 'center'}
#-- plot the generated trajectory x,y 
plot(x,y, main = "Generated data for trajectory",
     xlab = "Position x",
     ylab = "Position y")
```

Now that we generated noisy data for trajectory and velocity, we can sample the data $\theta_{1:n}^1$ and visualize them with a plot.

```{r, fig.align = 'center'}
theta_1 <- rep(0,k)
sigma <- 0.05

for (i in 1:k){
  theta_1[i] <- atan((y[i]-0.5)/(x[i]+1.5)) + rnorm(1, mean = 0, sd = sqrt(sigma))
}

plot(theta_1, main = "Sampled data for theta", ylab = "Value of theta")
```


# FUNCTIONS FOR LINEARIZATION

```{r}
##----------Function to compute C_k

C_k <-function(x,y,sx,sy){
  res<-rep(0,4)
  res[1]<- (sy-y)/((x-sx)^2+(y-sy)^2)
  res[2]<-(x-sx)/((x-sx)^2+(y-sy)^2)
  return(res)
}

##----------Function to compute D_k

D_k <-function(x,y,sx,sy){
  f<-atan((y-sy)/(x-sx))
  res<-f-x*(sy-y)/((x-sx)^2+(y-sy)^2)-
    y*(x-sx)/((x-sx)^2+(y-sy)^2)
  return(res)
}
```

# INITIALIZATION

```{r}
##-----------Initialization
A<-matrix(c(1,0,0,0,0,1,0,0,delta,0,1,0,0,delta,0,1),nrow = 4)
sx1<-(-1.5)
sy1<-0.5
X.HAT<-matrix(rep(NA,4*k),nrow=4)
P<-0.01*diag(1,4)
X.HAT[,1]<-mvrnorm(1,c(0,0,1,0),P)
```

# EXTENDED KALMAN FILTER

```{r}
for (t in 2:k){

  ##---------------Prediction Step
  Xmin<-A%*%X.HAT[,(t-1)]
  Pmin<-A%*%P%*%t(A)+Q
  C<-t(as.matrix(C_k(Xmin[1],Xmin[2],sx1,sy1)))
  D<-D_k(Xmin[1],Xmin[2],sx1,sy1)
  thetamin<-as.numeric(C%*%Xmin+D)

  ##--------------Correction Step
  K<-Pmin%*%t(C)%*%solve((C%*%Pmin%*%t(C)+sigma^2))
  X.HAT[,t]<-Xmin+K*(theta_1[t]-thetamin)
  P<-(diag(1,4)-K%*%C)%*%Pmin
}
```

# PLOTTING THE TRAJECTORIES

```{r}
x.hat<-X.HAT[1,]
y.hat<-X.HAT[2,]

df<-cbind.data.frame(Time=(1:k),x,y,x.hat,y.hat)
```

```{r}
library(ggplot2)
ggplot(data = df)+geom_point(aes(x=x,y=y,color="Real"))+
  geom_point(aes(x=x.hat,y=y.hat,color="Estimate"))+
  scale_color_manual(values=c("Real"="#E45C56","Estimate"="#566DE4"),
                     labels=c("Estimated trajectory","Real Trajectory")) + theme_bw()
```

# FUNCTIONS FOR LINEARIZATION // 2 SENSORS

```{r}
J_k2<-function(x,y,sx1,sy1,sx2,sy2){
  ##---------------------------------------Jacobian Matrix
  res<-matrix(rep(NA,4),nrow = 2)
  res[1,1]<-(sy1-y)/((x-sx1)^2+(y-sx1)^2)
  res[1,2]<-(x-sx1)/((x-sx1)^2+(y-sx1)^2)
  res[2,1]<-(sy2-y)/((x-sx2)^2+(y-sx2)^2)
  res[2,2]<-(x-sx2)/((x-sx2)^2+(y-sx2)^2)
  return(res)
}

C_k2<-function(x,y,sx1,sy1,sx2,sy2){
  ##-----------------------------------------Linear Function in the model
  res<-matrix(rep(0,8),nrow=2)
  res[,(1:2)]<-J_k2(x,y,sx1,sy1,sx2,sy2)
  return(res)
}

D_k2<-function(x,y,sx1,sy1,sx2,sy2){
  ##----------------------------------------Constant Part in the model
  f<-c(atan((y-sy1)/(x-sx1)),atan((y-sy2)/(x-sx2)))
  res<-f-J_k2(x,y,sx1,sy1,sx2,sy2)%*%as.matrix(c(x,y))
  return(res)
}
```

# NEW SENSORS DATA 

```{r}
sy2<-sx2<--1
R<-matrix(rep(NA,2*k),nrow = 2)

for (i in 1:k){
  R[,i]<-rnorm(2,0,sigma)
}

theta_2<--matrix(rep(NA,2*k),nrow = 2)

for (i in 1:k){
  theta_2[1,i]<-atan((y[i]-sy1)/(x[i]-sx1))
  theta_2[2,i]<-atan((y[i]-sy2)/(x[i]-sx2))
}

theta_2<-theta_2+R
```


# EXTENDED KALMAN FILTER

```{r}

##------------------Initialization

X.HAT<-matrix(rep(NA,4*k),nrow=4)
X.HAT[,1]<-mvrnorm(1,c(0,0,1,0),P)

for (t in 2:k){

  ##---------------Prediction Step
  Xmin<-A%*%X.HAT[,(t-1)]
  Pmin<-A%*%P%*%t(A)+Q
  C<-C_k2(Xmin[1],Xmin[2],sx1,sy1,sx2,sy2)
  D<-D_k2(Xmin[1],Xmin[2],sx1,sy1,sx2,sy2)
  thetamin<-C%*%Xmin+D
  ##--------------Corrction Step
  K<-Pmin%*%t(C)%*%solve((C%*%Pmin%*%t(C)+sigma^2*diag(1,2)))
  X.HAT[,t]<-Xmin+K%*%(matrix(theta_2[,t])-thetamin)
  P<-(diag(1,4)-K%*%C)%*%Pmin
}

```

# PLOTTING THE TRAJECTORIES

```{r}
x.hat<-X.HAT[1,]
y.hat<-X.HAT[2,]

df<-cbind.data.frame(Time=(1:k),x,y,x.hat,y.hat)

ggplot(data = df)+geom_point(aes(x=x,y=y,color="Real"))+
  geom_point(aes(x=x.hat,y=y.hat,color="Estimate"))+
  scale_color_manual(values=c("Real"="#E45C56","Estimate"="#566DE4"),
                     labels=c("Estimated trajectory","Real Trajectory")) + theme_bw()
```